\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-01-30 4-3-water-level.Rmd
matching example done :

The water level data comes from [Bureau of Meteorology](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) and has a copy in `weatherdata`. Here we extract the water course level and add a column annotate this data of type `river`. For the rainfall data, we will still use the `weatherdata::climate_full`, filtering for Victorian stations in 2020 should be pretty familiar by now. Again, we first look at where these stations are on the map first:

Now we use `match_sites()` to first pair the weather stations with the river gauges spatially and then apply the temporal matching on `prcp`. We will construct the interval based on peaks in `climate` since we would expect a lag effect for precipitation to flow into the river and cause a raise in river level, hence `temporal_independent = climate`. We select the 30 highest peak from the series to construct the match by setting `temporal_n_highest = 30`. This is a tuning parameter and you can start with 10% of the points of one series (here we have daily data for a year, 10% is roughly 30 points). `temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`.``

The output from temporal matching is also a cubble, with additional column `.dist` and `.group` inherent from spatial matching and `n_match` for the number of matched temporal peaks. Then you can use this output to plot the location of match or to look at the series:

```{r matching, out.width="100%", out.height="30%", fig.cap="asdfasd"}
knitr::include_graphics("figures/matching.png")
```

>>>>

Bureau of Meteorology also collects [water data](http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full) from river gauges, which includes electrical conductivity, turbidity, water course discharge, water course level, and water temperature. In particular, water level will interactive with precipitation from the climate data since rainfall will raise the water level in the river. Figure \ref{fig:matching-map} gives the location of available weather station and water gauges in Victoria.

```{r eval = FALSE}
climate <- weatherdata::climate_full %>%
  filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
  stretch() %>%
  filter(lubridate::year(date) == 2020) %>%
  tamp() %>%
  mutate(type = "climate")

river <- weatherdata::water %>%
  stretch() %>%
  select(date, Water_course_level) %>%
  rename(prcp = Water_course_level) %>% 
  tamp() %>%
  mutate(type = "river")
```

```{r matching-map, out.width="100%", fig.cap="Location of climate station and river gauges in Victoria, Australia."}
knitr::include_graphics("figures/matching-map.png")
```

From the map, a few water gauges and weather stations are close to each other and there is the potential that fluctuation of the water level can reflect the precipitation measured by the climate station. As introduced in Section 3.2, `match_sites()` can be used to match one source of data with another source in a cubble. Here `prcp` in `climate` will be matched to `Water_course_level` in `river`, which is specified in the argument `temporal_var_to_match`. `temporal_independent` controls the dataset that will be used to construct the interval (see the illustration in section 3.2 for details on this). Here the goal is to see if precipitation will be reflected by the water level in the river and hence a lagged effect of water level on precipitation. This will put precipitation data, `climate`, as the independent. Given there is one year worth of data, the number of peak (`temporal_n_highest`) to consider is slightly raised from a default 20 to 30. `temporal_min_match` filters out pairs don't have enough match and to return all the pairs, set `temporal_min_match` to `0`

```{r echo = TRUE}
res <- match_sites(river, climate,
                   temporal_var_to_match = prcp,
                   temporal_independent = climate,  
                   temporal_n_highest = 30,
                   temporal_min_match = 15)
```

The output from matching is also a cubble, with additional column `dist` and `group` produced from spatial matching and `n_match` from the temporal matching.

```{r}
res %>% print(n = 8)
```

Figure \ref{fig:matching} plots the matched pairs on the map or to view the matched series and with the chosen parameter, there are four pairs of matches, which all locates in the middle Victoria and the concurrent increase of precipitation and water level can be observed.

```{r matching, out.width="100%", fig.cap="Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The increases in the precipitation is reflected well by the water level."}
knitr::include_graphics("figures/matching.png")
```

### 2022-01-30 main.tex
matching example done :

Water data is collected by the
\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{Bureau
of Meteorology}. This includes \ldots{} water gauges, variables
collected. In particular, water level will interactive with the climate
variable precipitation as rains fall into the river will raise the water
level. Figure \ref{fig:matching-map} shows the available weather station
and water gauges in Victoria, Melbourne.

\begin{CodeChunk}
\begin{CodeInput}
R> climate <- weatherdata::climate_full %>%
+   filter(between(stringr::str_sub(id, 7, 8), 76, 90)) %>%
+   stretch() %>%
+   filter(lubridate::year(date) == 2020) %>%
+   tamp() %>%
+   mutate(type = "climate")
R> 
R> river <- weatherdata::water %>%
+   stretch() %>%
+   select(date, Water_course_level) %>%
+   rename(prcp = Water_course_level) %>% 
+   tamp() %>%
+   mutate(type = "river")
\end{CodeInput}
\end{CodeChunk}

@@

\caption[asdfasd]{asdfasd}\label{fig:matching-map}
\end{figure}
\end{CodeChunk}

From the map, a few water gauges and weather stations are close to each
other and we may be able to observe a spontaneous increase in the
precipitation and water level from the measures, potential with a little
delay in the river gauges. \texttt{match\_sites()} matches one source of
data with another source in both spatial and temporal dimension in a
cubble.

@@

The output from temporal matching is also a cubble, with additional
column \texttt{dist} and \texttt{group} produced from spatial matching
and \texttt{n\_match} from the number of peak matched temporally.

\begin{CodeChunk}
\begin{CodeOutput}
# cubble:   id [8]: nested form
# bbox:     [144.52, -37.73, 146.06, -36.55]
# temporal: date [date], prcp [dbl]
  id          name                lat  long ts         type   dist group n_match
  <chr>       <chr>             <dbl> <dbl> <list>     <chr> <dbl> <int>   <int>
1 405234      SEVEN CREEKS @ D~ -36.9  146. <tibble [~ river  6.15     5      21
2 ASN00082042 strathbogie       -36.8  146. <tibble [~ clim~  6.15     5      21
3 404207      HOLLAND CREEK @ ~ -36.6  146. <tibble [~ river  8.54    10      21
4 ASN00082170 benalla airport   -36.6  146. <tibble [~ clim~  8.54    10      21
5 230200      MARIBYRNONG RIVE~ -37.7  145. <tibble [~ river  6.17     6      19
# ... with 3 more rows
\end{CodeOutput}
\end{CodeChunk}

Then plots can be made to inspect the matched pairs on the map or to
view the co-movement of the two series:

@@

\caption[asdfasd]{asdfasd}\label{fig:matching}
\end{figure}
\end{CodeChunk}

With the chosen parameter, four pairs of stations stands out, which all
locates in the middle Victoria and the concurrent increase of
precipitation and water level can be observed.

>>>>

Bureau of Meteorology also collects
\href{http://www.bom.gov.au/metadata/catalogue/19115/ANZCW0503900528?template=full}{water
data} from river gauges, which includes electrical conductivity,
turbidity, water course discharge, water course level, and water
temperature. In particular, water level will interactive with
precipitation from the climate data since rainfall will raise the water
level in the river. Figure \ref{fig:matching-map} gives the location of
available weather station and water gauges in Victoria.

@@

\caption[Location of climate station and river gauges in Victoria, Australia]{Location of climate station and river gauges in Victoria, Australia.}\label{fig:matching-map}
\end{figure}
\end{CodeChunk}

From the map, a few water gauges and weather stations are close to each
other and there is the potential that fluctuation of the water level can
reflect the precipitation measured by the climate station. As introduced
in Section 3.2, \texttt{match\_sites()} can be used to match one source
of data with another source in a cubble. Here \texttt{prcp} in
\texttt{climate} will be matched to \texttt{Water\_course\_level} in
\texttt{river}, which is specified in the argument
\texttt{temporal\_var\_to\_match}. \texttt{temporal\_independent}
controls the dataset that will be used to construct the interval (see
the illustration in section 3.2 for details on this). Here the goal is
to see if precipitation will be reflected by the water level in the
river and hence a lagged effect of water level on precipitation. This
will put precipitation data, \texttt{climate}, as the independent. Given
there is one year worth of data, the number of peak
(\texttt{temporal\_n\_highest}) to consider is slightly raised from a
default 20 to 30. \texttt{temporal\_min\_match} filters out pairs don't
have enough match and to return all the pairs, set
\texttt{temporal\_min\_match} to \texttt{0}

@@

The output from matching is also a cubble, with additional column
\texttt{dist} and \texttt{group} produced from spatial matching and
\texttt{n\_match} from the temporal matching.

\begin{CodeChunk}
\begin{CodeOutput}
# cubble:   id [8]: nested form
# bbox:     [144.52, -37.73, 146.06, -36.55]
# temporal: date [date], prcp [dbl]
  id          name                lat  long ts         type   dist group n_match
  <chr>       <chr>             <dbl> <dbl> <list>     <chr> <dbl> <int>   <int>
1 405234      SEVEN CREEKS @ D~ -36.9  146. <tibble [~ river  6.15     5      21
2 ASN00082042 strathbogie       -36.8  146. <tibble [~ clim~  6.15     5      21
3 404207      HOLLAND CREEK @ ~ -36.6  146. <tibble [~ river  8.54    10      21
4 ASN00082170 benalla airport   -36.6  146. <tibble [~ clim~  8.54    10      21
5 230200      MARIBYRNONG RIVE~ -37.7  145. <tibble [~ river  6.17     6      19
6 ASN00086038 essendon airport  -37.7  145. <tibble [~ clim~  6.17     6      19
7 406213      CAMPASPE RIVER @~ -37.0  145. <tibble [~ river  1.84     1      18
8 ASN00088051 redesdale         -37.0  145. <tibble [~ clim~  1.84     1      18
\end{CodeOutput}
\end{CodeChunk}

Figure \ref{fig:matching} plots the matched pairs on the map or to view
the matched series and with the chosen parameter, there are four pairs
of matches, which all locates in the middle Victoria and the concurrent
increase of precipitation and water level can be observed.

@@

\caption[Matched weather stations and river gauges on the map (A) and across time (B)]{Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The increases in the precipitation is reflected well by the water level.}\label{fig:matching}
\end{figure}
\end{CodeChunk}
